---
title: "Analysis Data Preparation"
date: '`r format(Sys.time(), "%B %d, %Y")`'
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse=TRUE, prompt=TRUE, eval=TRUE, message=F, include=T,comment=NULL,fig.width = 8, warnings = FALSE, fig.height = 8,tidy.opts=list(width.cutoff=50),tidy=TRUE,cache = TRUE)
```

```{r packageCheck, include=FALSE}
mypacks <- c("tidyverse", "stringr", "data.table", "bit64")  # what packages are needed?
packs <- installed.packages()   # find installed package list
install.me <- mypacks[!(mypacks %in% packs[,"Package"])]  #what needs to be installed?
if (length(install.me) >= 1) install.packages(install.me, repos = "http://cran.us.r-project.org")   # install (if needed)
lapply(mypacks, library, character.only=TRUE)  # load all packages
```

# This file takes in raw orders data and conducts data manipulation. It then merges with network level data. 

# This file outputs data ready for analysis.

## Read raw orders data and network data
```{r}
orders <- fread("../../data/raw/orders_full.csv")
network <- fread("../../data/generated/retailer_network.csv")
```

## Manipulation of raw data

### get a list of unique retailers
```{r}
unique_retailers <- unique(orders[,list(RETAILER_NAME, RETAILER_ID, RETAILER_CATEGORY, RETAILER_PRICE_RANGE)])
```

### summarize the orders
```{r}
# Sales/orders amount
retailer_sales <- 
  orders[,
        .(sum(ORDER_TOTAL), .N),
        by = .(RETAILER_NAME)]
```

## Join, reorder, and write out

### joinwith network level and original retailer information
```{r}
out <- retailer_sales %>% 
  rename(RETAILER_TOTAL_SALES = V1,
         RETAILER_TOTAL_ORDERS = N) %>%
  left_join(unique_retailers, by = "RETAILER_NAME") %>%
  left_join(network, by = "RETAILER_NAME")
```

### Reorder columns
```{r}
out_final <- out %>% select(RETAILER_NAME, RETAILER_ID, RETAILER_CATEGORY,
                      RETAILER_PRICE_RANGE, RETAILER_TOTAL_SALES,
                      RETAILER_TOTAL_ORDERS, indegreeG:effectiveG
                      )
```


### Write out
```{r}
write.csv(out_final, "../../data/generated/data_for_analysis.csv", row.names = FALSE)
```