---
title: "Viz data Preparation"
date: '`r format(Sys.time(), "%B %d, %Y")`'
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse=TRUE, prompt=TRUE, eval=TRUE, message=F, include=T,comment=NULL,fig.width = 8, warnings = FALSE, fig.height = 8,tidy.opts=list(width.cutoff=50),tidy=TRUE,cache = TRUE)
```

```{r packageCheck, include=FALSE}
mypacks <- c("tidyverse", "data.table", "stringr")  # what packages are needed?
packs <- installed.packages()   # find installed package list
install.me <- mypacks[!(mypacks %in% packs[,"Package"])]  #what needs to be installed?
if (length(install.me) >= 1) install.packages(install.me, repos = "http://cran.us.r-project.org")   # install (if needed)
lapply(mypacks, library, character.only=TRUE)  # load all packages
```

# This file takes in raw data orders and conducts data manipulation. It then merges with network level orders and clustering results. 

# This file outputs retailer level data with all aggregated retailer info variables for visualization.

## Read raw data orders
```{r}
orders <- fread("../../data/raw/orders_full.csv")
network <- fread("../../data/generated/retailer_network.csv")
clustering <- fread("../../data/generated/clustering_results.csv")
```

## Manipulation of raw orders data

### get a list of unique retailers
```{r}
unique_retailers <- unique(orders[,list(RETAILER_NAME, RETAILER_CATEGORY, RETAILER_PRICE_RANGE)])
```


### Create order month
```{r}
orders <- orders[, ORDER_MONTH := month(orders$ORDER_DATE)]
```

### group by (create new vars first, dt command)
```{r}
# GENDER, IS_CROSS_SELL, month
temp_1 <- 
  orders[,
        .(sum(ORDER_TOTAL), .N),
        by = .(RETAILER_NAME, MEMBER_GENDER, ORDER_MONTH, IS_CROSS_SELL)]

# MEMBER_ID, FIRST VISIT RETAILER
temp <- unique(orders[,list(MEMBER_ID, MEMBER_FIRST_VISIT_RETAILER)])

temp_fr <- temp[,
        .N,
        by = .(MEMBER_FIRST_VISIT_RETAILER)]
```

#### prepare new info ready for join (dplyr command)
```{r}
temp_1 <- temp_1 %>%
  rename(RETAILER_TOTAL_SALES = V1,
         RETAILER_TOTAL_ORDERS = N)

temp_fr <- temp_fr %>%
  rename(RETAILER_NUM_FIRST_VISIT_CUST_YRLY = N,
         RETAILER_NAME = MEMBER_FIRST_VISIT_RETAILER)
```

## Join, reorder, and write out

### Join with network level data and clustering results
```{r}
out <- temp_1 %>% 
  left_join(temp_fr, by = "RETAILER_NAME") %>%
  left_join(network, by = "RETAILER_NAME") %>%
  left_join(clustering, by = "RETAILER_NAME") %>%
  left_join(unique_retailers, by = "RETAILER_NAME") %>%
  arrange(RETAILER_NAME, ORDER_MONTH, MEMBER_GENDER, IS_CROSS_SELL)
```

### Reorder
```{r}
out_final <- out %>% 
  select(RETAILER_NAME, RETAILER_ID, RETAILER_CATEGORY, RETAILER_PRICE_RANGE, 
         RETAILER_CLUSTER_NUMBER, RETAILER_CLUSTER_NAME,
         RETAILER_NUM_FIRST_VISIT_CUST_YRLY,
         ORDER_MONTH, MEMBER_GENDER, IS_CROSS_SELL,
         RETAILER_TOTAL_SALES, RETAILER_TOTAL_ORDERS, indegreeG:effectiveG)
```

### Write out
```{r}
write.csv(out_final, "../../data/generated/data_for_viz.csv", row.names = FALSE)
```